Two measures are better than one: Combining iconicity ratings and guessing experiments for a more nuanced picture of iconicity in the lexicon

### 2022-02-28 paper.Rmd
deleted iconicity box code
anonymise

Behavioural approaches involve the collection of data from naive experimental participants, which can reveal something about the iconicity of given form-meaning pairings. The only requirement is the human participants for the study, and the largest amount of iconicity measurements from the widest variety of languages have been collected using these methods. There are two main approaches: iconicity rating tasks and guessing experiments.

@@

For the new version of the guessing experiments, we minimised the amount of English words involved in the task by having participants match a single English translation to the corresponding word in Japanese, choosing between two possible Japanese words. The instructions used were as follows:

@@

Since /a/ is often the most phonologically distant vowel from many of the vowels in the Japanese sound system, to avoid the foil words all having similar vowels, a different approach was used for substituting vowels. Instead of using feature values, vowels were substituted with the vowel obtained by rotating the vowel space 180 clockwise or counterclockwise. So /u/ is substituted with either /a/ or /i/, and /a/ is substituted with either /i/ or /u/ (see Figure \@ref(fig:vowels)). Finally, reduplicated words were ‘unreduplicated’ when creating foil words. For example, the foils for *fuwafuwa* were *watʃi*, *nati* and *ridːu* (not *watʃiwatʃi*, *natinati* and *ridːuridːu*).

@@

Focusing first on the data from the current study, we can see that the guesses provide a more even spread of measures compared to the ratings. Between the two guessing methods, the method of guessing between words provides a wider spread of measures than guessing between translations--both at the very high end of the scale (highly guessable words) and at the very low end of the scale (poorly guessable words). If we look at the ratings from native Japanese speakers, these have a pretty equivalent spread to the guesses between words. Their distribution is slightly narrower at the very top end of the scale. However, the data from the current study was chosen from a domain known to be high in iconicity, whereas the data from Thompson et al. was more varied, with proportionally less ideophones. So, this could simply be a reflection of the differences between the two datasets rather than relating to the methods used.

@@

However, ideophones were consistently rated *higher* in iconicity than prosaic words—even when guessed at the same accuracies. We believe this is due to the majority of ideophones in the dataset being reduplicated. The two lowest rated ideophones–*sappari* ‘clean’ and *guttari ‘*exhausted*’*–were among the few in the data that do not use a reduplicated template. Previous work has found a relationship between iconicity ratings and structural markedness [@dingemanse_playful_2020]. Ordinarily when we use words, the form of the words is not important–only their content. However, when iconicity is involved the form of the words *is* important, and structural markedness may serve as a metacommunicative cue to signal this. The results from the rating task suggest that structural markedness (in this case, reduplication) leads to the perception of iconicity. Interestingly, however, it does not seem to confer an advantage–over and above that provided by associations between form and meaning–in the guessing task. What the ratings seem to be picking up on is both form-meaning associations *and* structural markedness. If they were only picking up on structural markedness, then we would not expect the ratings to vary with the guesses in the same way for both ideophones and prosaic words. However, if they were only picking up on form-meaning associations (like the guesses seem to), then we would not expect ideophones to consistently receive higher iconicity ratings, even when guessed at the same accuracies.

@@

It is interesting that conflicting associations appear to have a stronger effect in ideophones (some of which were guessed at accuracies significantly below chance) than in prosaic words (which were guessed below chance only rarely). Perhaps because participants are more likely to view form-meaning relationships in ideophones as meaningful (due to their structural markedness), they are more likely to attach conflicting associations to ideophones, leading to a stronger sense that the ideophone should correspond to the *opposite* of the given meaning [for more on this see discussion in @webster_2017; @akita_ideophone_2019: 19]. This is supported by the larger person-total correlations reported for ideophones compared to prosaic words in Section 3.3; that participants were more consistent with each other when rating ideophones suggests that intuitions about these words were more pervasive.

>>>>

Behavioural approaches involve the collection of data from naive experimental participants, which can reveal something about the iconicity of given form-meaning pairings. The only requirement is the human participants for the study, and the largest number of iconicity measurements from the widest variety of languages have been collected using these methods. There are two main approaches: iconicity rating tasks and guessing experiments.

@@

For the new version of the guessing experiments, we minimised the number of English words involved in the task by having participants match a single English translation to the corresponding word in Japanese, choosing between two possible Japanese words. The instructions used were as follows:

@@

Since /a/ is often the most phonologically distant vowel from many of the vowels in the Japanese sound system, to avoid the foil words all having similar vowels, a different approach was used for substituting vowels. Instead of using feature values, vowels were substituted with the vowel obtained by rotating the vowels at the endpoints of the vowel space triangle 120 degrees clockwise or counterclockwise, and flipping the vowels in the middle. So /u/ is substituted with either /a/ or /i/, /a/ is substituted with either /i/ or /u/, /i/ is substituted with either /a/ or /u/, and /e/ and /o/ are substituted with each other (see Figure \@ref(fig:vowels)). Finally, reduplicated words were ‘unreduplicated’ when creating foil words. For example, the foils for *fuwafuwa* were *watʃi*, *nati* and *ridːu* (not *watʃiwatʃi*, *natinati* and *ridːuridːu*).

@@

Focusing first on the data from the current study, we can see that the guesses provide a more even spread of measures compared to the ratings. Between the two guessing methods, the method of guessing between words provides a wider spread of measures than guessing between translations--both at the very high end of the scale (highly guessable words) and at the very low end of the scale (poorly guessable words). If we look at the ratings from native Japanese speakers, these have a pretty equivalent spread to the guesses between words. Their distribution is slightly narrower at the very top end of the scale. However, the data from the current study was chosen from a domain known to be high in iconicity, whereas the data from Thompson et al. was more varied, with proportionally fewer ideophones. So, this could simply be a reflection of the differences between the two datasets rather than relating to the methods used.

@@

However, ideophones were consistently rated *higher* in iconicity than prosaic words—even when guessed at the same accuracies. We believe this is due to the majority of ideophones in the dataset being reduplicated. The two lowest rated ideophones–*sappari* ‘clean’ and *guttari ‘*exhausted*’*–were among the few in the data that do not use a reduplicated template. Previous work has found a relationship between iconicity ratings and structural markedness [@dingemanse_playful_2020]. Ordinarily when we use words, the form of the words is not important–only their content. However, when iconicity is involved the form of the words *is* important, and structural markedness may serve as a metacommunicative cue to signal this. The results from the rating task suggest that structural markedness (in this case, reduplication) leads to the perception of iconicity. Interestingly, however, it does not seem to confer an advantage–over and above that provided by associations between form and meaning–in the guessing task. What the ratings seem to be picking up on is both form-meaning associations *and* structural markedness. If they were only picking up on structural markedness, then we would not expect the ratings to correspond with the guesses in the same way for both ideophones and prosaic words. However, if they were only picking up on form-meaning associations (like the guesses seem to), then we would not expect ideophones to consistently receive higher iconicity ratings, even when guessed at the same accuracies.

@@

It is interesting that conflicting associations appear to have a stronger effect in ideophones (some of which were guessed at accuracies significantly below chance, reflecting a strong bias *against* the word having the specified meaning) than in prosaic words (which were guessed below chance only rarely). Perhaps because participants are more likely to view form-meaning relationships in ideophones as meaningful (due to their structural markedness), they attach more weight to perceived conflicts between a form and its given meaning, resulting in a stronger pull away from the correct form [for more on this see discussion in @webster_2017; @akita_ideophone_2019: 19]. This is supported by the larger person-total correlations reported for ideophones compared to prosaic words in Section 3.3; that participants were more consistent with each other when rating ideophones suggests that intuitions about these words were more pervasive.
